Technique for \earlywarning\ filtering for compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-29 introduction.tex
In section 1, paragraph 1, switched 'discuss' with 'present'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1181 ccdd05d5-56c1-461a-bf98-85f6876523b4

As a compact binary system loses energy to gravitational waves (\GW{}s), its
orbital separation decays, leading to a runaway inspiral with the \GW{}
amplitude and frequency increasing until the system eventually merges.
If a neutron star (\textsc{ns}) is involved, it may become tidally disrupted near
the merger and fuel a bright electromagnetic (\EM{}) counterpart
\citep{shibata:2007}.  The \GW{} signal is in principle detectable
before merger.  Effort from both the \GW{} and astronomy communities may make
it possible to use \GW{} observations as an early warning trigger for \EM{}
followup. In the first generation of ground-based laser interferometers, the \GW{}
community initiated a project to send alerts when potential \GW{} transients
were observed in order to trigger followup observations by \EM{} telescopes.
The typical latencies were 30 minutes \citep{HugheyGWPAW2011}, which was an
important achievement, but too late to catch any prompt \EM{}
emission. We have the ambition of reporting \GW{} candidates not minutes after
the merger, as has already been accomplished, but seconds before.  We discuss
one essential ingredient in this effort, a computationally inexpensive
low-latency filtering algorithm for detecting inspiral signals in \GW{}
data.  We also present the prospects for advanced \GW\ detectors and
discuss other areas of work that would be required for rapid analysis.

>>>>

As a compact binary system loses energy to gravitational waves (\GW{}s), its
orbital separation decays, leading to a runaway inspiral with the \GW{}
amplitude and frequency increasing until the system eventually merges.
If a neutron star (\textsc{ns}) is involved, it may become tidally disrupted near
the merger and fuel a bright electromagnetic (\EM{}) counterpart
\citep{shibata:2007}.  The \GW{} signal is in principle detectable
before merger.  Effort from both the \GW{} and astronomy communities may make
it possible to use \GW{} observations as an early warning trigger for \EM{}
followup. In the first generation of ground-based laser interferometers, the \GW{}
community initiated a project to send alerts when potential \GW{} transients
were observed in order to trigger followup observations by \EM{} telescopes.
The typical latencies were 30 minutes \citep{HugheyGWPAW2011}, which was an
important achievement, but too late to catch any prompt \EM{}
emission. We have the ambition of reporting \GW{} candidates not minutes after
the merger, as has already been accomplished, but seconds before.  We present
one essential ingredient in this effort, a computationally inexpensive
low-latency filtering algorithm for detecting inspiral signals in \GW{}
data.  We also discuss the prospects for advanced \GW\ detectors and
discuss other areas of work that would be required for rapid analysis.
